Taking Linear Logic Apart

### 2018-04-25 doc/tlla2018/main.tex
a bit of intro

\begin{abstract}
\end{abstract}

>>>>

\begin{abstract}
We introduce the calculus of Hypersequent Classical Processes (\hcp), which extends the research line on the relationship between linear logic and bridges the gap between the structures of operators used as proof terms in previous work and those of the standard $\pi$-calculus (in particular parallel operator and restriction).
The key twist from previous work is registering parallelism in typing judgements, by generalising linear logic judgements from one sequents to many (hypersequents).
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Classical Processes (CP) \cite{wadler2014} is a process calculus inspired by the correspondence between the session-typed $\pi$-calculus and linear logic \cite{caires2010}, where processes correspond to proofs, session types (communication protocols) to propositions, and communication to cut elimination. Bridging process languages to linear logic paves the way to apply methods developed in one field to the other. This already worked for a few results, in both directions. For example, the proof theory of linear logic can be used to guarantee progress for processes \cite{caires2010,wadler2014}, and multiparty session types, originally developed for processes \cite{honda2016}, inspired a generalisation of the standard cut rule to the composition of an arbitrary number of proofs, allowing for safe circular dependencies among proofs \cite{carbone17}.

The hallmark of CP is that the semantics of processes is given by sound proof transformations in Classical Linear Logic (CLL). While this permits reusing the metatheory of linear logic ``as is'' to reason about process behaviour (\eg, cut elimination yields communication progress), it also exhibits some fundamental discrepancies with the key operators of the $\pi$-calculus \cite{milner1992a,milner1992b}.

Some discrepancies are syntactic. For example, the term for output of a linear name is $\cpSend{x}{y}{P}{Q}$, read ``send $y$ over $x$ and proceed as $P$ in parallel to $Q$''. Notice that the term constructor for output here actually 
takes $x$, $y$, $P$ and $Q$ as parameters at the same time. This discrepancy is caused by adopting processes as proof terms for CLL: the typing rule for 
output (\ie, the $\tens$ rule of CLL) checks that the processes respectively implementing the behaviours of $y$ ($P$) and of $x$ ($Q$) share no resources, by taking two premises ($P$ and $Q$).
In general, there is no independent parallel term $P \mid Q$ in the grammar of CP, and even if we added it as the $\NOM{Mix}$ rule suggested in the original presentation of CP \cite{wadler2014}, it would not allow $P$ and $Q$ to communicate as in standard $\pi$-calculus. Synchronisation is governed instead by the restriction operator $\piNew{x}{(\piPar{P}{Q})}$ (we use the latest syntax for CP, from \cite{carbone2016}), which links $x$ at $P$ with $y$ at $Q$ to enable communication. Again, parallel is mixed with another operator (restriction here), but in this case it means that $P$ and $Q$ will communicate

The discrepancies carry over from syntax (and typing) to semantics.
The rule for reducing an output with an input in CP is the following.
\[
	\cpCut{x}{\cpSend{x}{y}{P}{Q}}{\cpRecv{x}{y}{R}}
  \Longrightarrow
  \cpCut{y}{P}{\cpCut{x}{Q}{R}}
\]
Notice how the rule needs to inspect the structure of the continuation of the output term ($\piPar{P}{Q}$) to produce a typable structure for the resulting network, by nesting restrictions appropriately.
